% !TeX root = ./main.tex
\documentclass[3p]{elsarticle}

\usepackage{paralist}
\usepackage{lineno,hyperref}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{hyperref}
\usepackage{paralist}
\usepackage{ifthen}
%\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{amsmath,amssymb}
% \usepackage{algpseudocode}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{soul}


\usepackage{lscape}


\definecolor{vuborange}{rgb}{1.0,0.40,0.0}


\newcommand\m[1]{\ensuremath{#1}\xspace}
\newcommand\ltrue{\m{\mathbf{t}}}
\newcommand\lunkn{\m{\mathbf{u}}}
\newcommand\lfalse{\m{\mathbf{f}}}
\newcommand\leqp{\m{\leq_p}}
\newcommand\geqp{\m{\geq_p}}
\newcommand\entails{\m{\models}}
% \newcommand\land{\m{\wedge}}
\newcommand\limplies{\m{\Rightarrow}}
\newcommand\ourtool{\textsc{ZebraTutor}\xspace}
\newcommand\idp{\textsc{IDP}\xspace}
\usepackage{amsthm}
\usepackage{thmtools} 

\newtheorem{thm}{Theorem}
\newtheorem{definition}[thm]{Definition}
\newtheorem{prop}{Property}
\newtheorem{property}[prop]{Property}
\newtheorem{ex}{Example}
\newtheorem{example}[ex]{Example}
\newcommand\xxx{\m{\overline{x}}}
\newcommand\ddd{\m{\overline{d}}}

%To easily allow for saving space
\newcommand{\myparagraph}[1]{\subsection{#1}}

\newcommand\allconstraints{\m{T}}
\modulolinenumbers[5]

\journal{Journal of Artificial Intelligence}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%
%% `User comments`
\newcommand\comment[1]{\marginpar{\tiny #1}}
\renewcommand\comment[1]{#1}
\newcommand{\bart}[1]{{\comment{\color{red}\textsc{BB:}#1}}}
\newcommand{\tias}[1]{{\comment{\color{blue}\textsc{TG:}#1}}}
\newcommand{\emilio}[1]{{\comment{\color{red}\textsc{EG:}#1}}}
\newcommand{\todelete}[1]{{\comment{\color{red}\st{#1}}}}
\newcommand{\todo}[1]{{\comment{\color{red}\textsc{TODO:}#1}}}
\newcommand\ignore[1]{}
\newcommand{\toprule}{}
\newcommand{\midrule}{\hline}
\newcommand{\bottomrule}{}

% comment out for removing the comments
% \renewcommand\comment[1]{}

%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}


\begin{frontmatter}

\title{A framework for step-wise explaining how to solve constraint satisfaction problems}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author[mymainaddress]{Bart Bogaerts, Emilio Gamba, Tias Guns}
\address{Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussel, Belgium}
\ead{\{firstname.lastname\}@vub.be}

\begin{abstract}
We explore the problem of step-wise explaining how to solve constraint satisfaction problems, with a use case on logic grid puzzles.
More specifically, we study the problem of explaining the inference steps that one can take during propagation, in a way that is easy to interpret for a person.
Thereby, we aim to give the constraint solver explainable agency, which can help in building trust in the solver by being able to understand and even learn from the explanations.
The main challenge is that of finding a sequence of \textit{simple} explanations, where each explanation should aim to be as cognitively easy as possible for a human to verify and understand. 
This contrasts with the arbitrary combination of facts and constraints that the solver may use when propagating.
We propose the use of a cost function to quantify how simple an individual explanation of an inference step is, and identify the explanation-production problem of finding the best sequence of explanations of a CSP. 
Our approach is agnostic of the underlying constraint propagation mechanisms, and can provide explanations even for inference steps resulting from combinations of constraints. 
In case multiple constraints are involved, we also develop a mechanism that allows to break the most difficult steps up and thus gives the user the ability to \emph{zoom in} on specific parts of the explanation. 
Our proposed algorithm iteratively constructs the explanation sequence by using an optimistic estimate of the cost function to guide the search for the best explanation at each step.
Our experiments on logic grid puzzles show the feasibility of the approach in terms of the quality of the individual explanations and the resulting explanation sequences obtained.
\end{abstract}

\begin{keyword}
Artificial Intelligence \sep Constraint Solving \sep Explanation
%\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
%\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}\label{sec:intro}
\input{files/1_intro}

\section{Related work}\label{sec:related-work}
\input{files/2_related_work}

\section{Background}\label{sec:background}\label{sec:prelims}
\input{files/2_background}

\section{Problem definition}\label{sec:problem-definition}
\input{files/3_problem_definition}

\section{Nested Explanations}\label{sec:nested-explanation}
\input{files/4_nested_explanations}

\section{Explanation-producing search}\label{sec:expl-gen-prod}
\input{files/4_decomposition_prop_unsat}


\section{Explanations for logic grid puzzles}\label{sec:zebra}
\input{files/5_good_ordering}

\section{Logic grid puzzles: From natural language clues to typed first-order logic}\label{sec:holistic}
\input{files/6_pipeline}

\section{Experiments}\label{sec:experiments}
\input{files/7_experiments}

\section{Discussion, future work, and conclusions}\label{sec:conclusion}
\input{files/9_concl}

% \section*{Author Contributions}
% \bart{Is this required? Where does this come from? 

% If it is, I like the second style more. Also, We should Acknowledge Jens's work indeed, but probably rather in the Acknowledgement section}


% \todo{Please add anything relevant, i don't remember who did what specifically}
% \emilio{e.g All authors contributed equally to the writing of the paper?}

% \noindent\textbf{Bart Bogaerts} helped with the \textit{Software}, the \textit{Visualisation} and \textit{Writing} of the paper.

% \noindent\textbf{Emilio Gamba} helped with the \textit{Visualisation}, the \textit{Experiments}, and the \textit{Writing} of the paper.

% \noindent\textbf{Tias Guns} helped with \textit{Conceptualisation} of the research and experiments, \textit{Supervision} and the \textit{Writing} of the paper.

% \tias{or something like (maybe Bart wants to add something more? smth like grateful to Jens for system describe din Sec 8?):} BB and TG conceived the research, BB and EG implemented the software, EG performed the experiments and all authors contributed equally to the design of the experiments and the writing of the paper. \tias{First time I have to do this}

\section*{Acknowledgements}
This research received funding from the Flemish Government under the ``Onderzoeksprogramma Artifici\"ele Intelligentie (AI) Vlaanderen'' programme. 
We thank Jens Claes (and his master thesis supervisor Marc Denecker) for the implementation of a typed extension of the Blackburn \& Bos framework as part of his master's thesis, as well as Rocsildes Canoy for his help with the NLP aspect of the information pipeline. %\todo{Thank Jens (and Rocs?) (and co-supervisors of Jens' thesis?)}

% \section*{References}

\bibliography{mybibfile,ref,krrlib}

\end{document}
