% !TeX root = ./main.tex
We presented a \hitsetbased algorithm for finding \textit{optimal constrained} unsatisfiable subsets, with an application in generating explanation sequence for constraint satisfaction problems. Constrainedness %\deleted{, incrementality}
and a domain-specific \grow method were key to generating such sequences in a reasonable amount of time.
Incrementality still remains an unresolved question especially when considering applications where incrementality can also play across different queries. \bart{what does this sentence mean? It remains an unresolved question... WHat is unresolved? Why do we only mention this now? We have been pretending to resolve it the entire time. 
And what does ``where incrementality can also play accross different queries'' come from. It has been playing accross different queries (each explanation stap can be seen as a different query) the entire time}

%Summarised, while \citet{ecai/BogaertsGCG20} had to generate many MUSs for finding a single next explanation step, the step-wise explanation generation directly benefits from the \comus approach of finding the optimal explanation candidate at any point in the sequence.

%There are many trade-offs to be made in the algorithm, including time spent computing optimal or non-optimal hitting sets, how to reuse information and how to grow the satisfiable subsets. While we studied the key algorithmic dimensions of information reuse, a deeper study of alternative approaches to balance these trade-offs is needed, for instance comparing different methods to grow a satisfiable subset, and balancing the trade-off between finding hitting sets of high quality versus having to compute fewer hitting sets.

With the observed impact of different `\grow' methods, an open question remains whether we can quantify precisely and in a generic way what a \textit{good} or even the best set-to-hit is in a hitting set approach. 
While we focused on generating entire sequences, another question could be what the best method is for finding a single explanation step, that is a single OCUS.   \bart{what does this last sentence mean? All our algorithms aim at finding a single OCUS. The only bit of sequence focus we have is on incrementality, but that is a minor part of the paper...}

Our experimental results show that the majority of OCUS calls can be performed relatively fast, and hence are encouraging for (near) real-time settings such as interactive configuration applications~\cite{van2017kb}.

%\deleted{This would be important for example in interactive configuration applications ; where incrementality can also play across different queries.}

The synergies of our approach with the more general problem of QMaxSAT \cite{DBLP:journals/constraints/IgnatievJM16} is another open question.

%The goal of explaining satisfaction problems fits in wider goal of human-machine decision making, which is related to interactive constraint solving~\cite{putnam2019toward}. The use of MUSs in this context has been explored by . For interactivity, finding a single O(C)US fast would be the focus, where incrementality can play a big role.

The concept of OUS, incremental OUS and OCUS are not limited to explanations of satisfaction problems and we are keen to explore other applications too.
A general direction here are \textit{optimisation} problems and the role of the objective function in explanations.
%From the explanation point of view, a next challenge is how to compute explanations for optimisation problems, that is, where decisions are made based on search and not just propagation. We believe a constrained OUS algorithm can also play a key part in that. Finally, an open challenge is that of defining appropriate cost functions for generating explanations, including how to evaluate what a ``good'' explanation sequence is. While we currently find a sequence where each step is optimal with respect to the cost function, we have yet to consider whether it is possible to optimize a cost function over the entire sequence as well.
%\deleted{Within satisfaction, both MUSs and \hitsetbased algorithms are also investigated in the context of explaining machine learning decisions~\cite{ignatiev2019abduction}. 
%In this context, an open direction for future work is to examine the potential benefit of using optimal MUS search and whether constrainedness properties can further boost its possibilities.}
%
%\emilio{Recently, MUSs have been used to formally define abduction-based explanations for machine learning classifiers \cite{ignatiev2019abduction}.
%	In that work, from a given classifier, and a concrete classified instance, a propositional theory explains which set of feature values of the instance are responsible for the predicted class.
%	We see opportunities to also use OCUS in this setting.
%	For instance, OCUS allows the formulation of structural constraints on the explanations, and (ii) the quantification of explanation difficulty using the optimality criterion.}
%\begin{itemize}
%    \item Conclusion on resutls
%    \item Challenges next : extension to XOPT ? 
%    \item Applicability on other problems ?
%    \item Characterizing explanation difficulty
%\end{itemize}

% \paragraph{Acknowledgments}
% \textit{This research received partial funding from the Flemish Government (AI Research Program) and the FWO Flanders project G070521N.}
% \bart{Why is this in a paragaraph environment (not sectionstar) and why italic? Is this this style guidelines or simply fantasy? I would write: }
\section*{Acknowledgment}
This research received partial funding
from the Flemish Government (AI Research Program) and the
FWO Flanders project G070521N.
